---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  word_document: default
---


## Loading and preprocessing the data
The file is first downloaded, unzipped, and read into R as a CSV file.
```{r,echo=TRUE}
#download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",destfile = "Rmark",method = "curl")
unzip("Rmark")
act = read.csv("activity.csv")
```
## What is mean total number of steps taken per day?
We use tapply to find the total of steps taken each day. This value is shown as gh.
```{r}
gh = tapply(act[,1],act[,2],sum,na.rm = T)
hg = as.data.frame(gh)
hg
```
We now can make a histogram of hg to show the distribution of the frequencies of the daily averages.
```{r}
hist(hg[,1],main = "Historgam of Daily Averages", xlab = "Number of Steps Taken per Day",ylab = "Frequency")
```

Lastly we can calculate the the mean and median of our daily averages.

```{r}
mean(hg[,1])
median(hg[,1])
```
## What is the average daily activity pattern?
We begin my taking the average by time of day.
```{r,echo=T}
ip = tapply(act[,1],act[,3],mean,na.rm = T)
```

Now we can plot our graph of average numbers of steps taken by hour of the day.

```{r}
up = as.data.frame(ip)
plot(unique(act[,3]),ip,type = 'l',main = "Average number of Steps by Time of day",xlab = "Hour of the Day",ylab = "Number of Steps")
```
Now we need to determine how at what time and how many steps was the maximum value. This is shown below.
```{r}
up[up[,1]==max(ip),]
```


## Imputing missing values
Here we are counting the total amount of NA values in the dataframe. 
```{r,echo=T}
Total.NA = sum(is.na(act[,1]))
```

The strategy for removing NA values will be to replace them with the total daily median value for that time interval.

```{r}
act1 = act
up[,2] = row.names(up)
for(i in 1: nrow(act1)){
  if(is.na(act1[i,1])){
    act1[i,1] = up[up[,2] == act[i,3],][1]
  }else{
    act1[i,1] = act1[i,1]
  }
}
```
Historgram of new data
```{r}

gh1 = tapply(act1[,1],act1[,2],sum)
hg1 = as.data.frame(gh1)

hist(hg1[,1],main = "Histogram of Steps Taken per Day with NA Removed",xlab = "Number of steps taken per day", ylab = "Frequency")

```
Mean and Median of new data with NAs removed.

```{r}
mean(as.numeric(act1[,1]))
median(as.numeric(act1[,1]))
```
## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(lattice)
act1$day <- weekdays(as.Date(act1[,2]))
for( i in 1:nrow(act1))
if(act1[i,"day"] %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")){
  act1[i,"day"]= "Weekday"
}else{
  act1[i,"day"] = "Weekend"
}
sub.weekday = act1[act1[,"day"]=="Weekday",]
sub.weekend = act1[act1[,"day"] == "Weekend",]
wkd = tapply(sub.weekday[,1],sub.weekday[,3],mean)
wke = tapply(sub.weekend[,1],sub.weekend[,3],mean)
wkd = as.data.frame(wkd)
wkd$day = "weekday"
wkd$time = row.names(wkd)
wke = as.data.frame(wke)
wke$day = "weekend"
wke$time = row.names(wke)
colnames(wke) = c("wkd","day","time")
total = rbind(wke,wkd) 
xyplot(total[,1]~as.numeric(total[,3])|total[,2],layout = c(1,3),xlab = "Time of Day",ylab = "Number of Steps",type = 'l')
```
